### 2008-01-25 docs/motivation/motivation.tex
Initial creation of the motivating paper.

\documentclass[conference]{IEEEtran}

\usepackage[noadjust]{cite}

%\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
\usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
%\fi

\usepackage{url}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{threeparttable}
\usepackage{multirow}

\begin{document}

\title{Using Altruistic Peers to Supplement Downloading Bandwidth}
\author{\IEEEauthorblockN{Cameron Dale}
\IEEEauthorblockA{School of Computing Science\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: camerond@cs.sfu.ca}}

\maketitle

\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{related}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background}
\label{background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  Section  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{conclusions}

\bibliographystyle{IEEEtran}
\bibliography{./IEEEabrv,./all}

\end{document}